Uncertainty Quantification in Multiscale Models of Charge Transport in Organic Semiconductors: Influence of the Exchange-Correlation Functional

### 2025-05-21 submit/paper.tex
Checked reply and manuscript. Few suggestions in red and a first attempt at reply to point 12.

Uncertainty quantification (UQ) is concept from computational science which allows for an estimation of confidence intervals for a quantity of interest (QoI) and an analysis of its sensitivity in models of a physical system that contain uncertain, maybe empirical, or noisy, parameters~\cite{sarkar_uncertainty_2017,oconnor_quantifying_2024, chernatynskiy_uncertainty_2013, suleimenova_tutorial_2021,coveney_reliability_2021, coveney_when_2021}. Many common UQ studies focus on models with (partial) differential equations, e.g., drift-diffusion equations in which the diffusivity as parameter, and assume a certain distribution for the values of the parameter(s). For the multiscale model of charge transport, it is not straightforward to cast the large variety of available exchange-correlation functionals into the role of a model parameter with some distribution. To keep the problem tractable, we focus here instead specifically on the exchange part in hybrid functionals~\cite{perdew_rationale_1996,marques_densitybased_2011}, in which a DFT model for the exchange is mixed using a weighting factor $\alpha_\text{HFX}$ with a Hartree--Fock type exchange, i.e.,
%
\begin{equation}
  E_\text{x} = \alpha_\text{HFX} E_\text{x}^\text{HF} + (1-\alpha_\text{HFX})E_\text{x}^\text{DFT}.
  \label{equ:hybrid}
\end{equation}
%
\change{This choice is motivated by the observation that the simulated charge-carrier time-of-flight (ToF) for a set of commonly used functionals --- PBE~\cite{PhysRevLett.77.3865}, B3LYP~\cite{10.1063/1.464913}, PBE0~\cite{adamo_toward_1999}, BHLYP~\cite{10.1063/1.464913,lee_development_1988}, M06L~\cite{10.1063/1.2370993}, TPSS~\cite{PhysRevLett.91.146401} --- (see Tab.~\ref{tab:actual_functionals} in the Appendix) is more sensitive to the amount of included Hartree--Fock exchange than to variations in the generalized-gradient approximation (GGA) parametrization. While these functionals vary in more details, the data shows that for functionals with varying \ahfx, the simulated ToF spans two orders of magnitude, whereas among non-hybrid (meta-)GGA functionals the ToF results on the same order of magnitude.} More specifically, we take as the basis the PBE0 functional and scrutinize (i) how the predictions of the multiscale model of charge transport are affected by variation of $\alpha_\text{HFX}$ as a proxy for uncertainties in the choice of DFT functionals, (ii) what the level of confidence is in quantitative predictions, and (iii) what are the most sensitive quantities in the model. In this sense, Eq.~\ref{equ:hybrid} is deceptively simple. For each value of $\alpha_\text{HFX}$, the graph $G$ is constructed using the respective value of the reorganization energy, the $N_\text{mol}$ site energies, and the $N_\text{pair}$ coupling elements, and the dimensionality of the problem from the perspective of UQ is $N_\text{UQ}=1+N_\text{mol}+N_\text{pair}$, which can easily be on the order of $10^{4}-10^{5}$. We consider the simulated time-of-flight (ToF) and the associated mobility the QoI in the following (see Fig.~\ref{fig:MSM}(c)) which then are subject to uncertainties in these $N_\text{UQ}$ parameters, {\em stemming} from the variation in $\alpha_\text{HFX}$. As a prototypical system, we will study hole transport in an amorphous morphology of 2-methyl-9, 10-bis(naphthalen-2-yl)anthracene (MADN), a wide-gap semiconductor that is used extensively as an ambipolar host material in organic light-emitting diodes~\cite{ko_accurate_2019, chang_great_2017}.

@@

The four respective distributions of $\tau$ in the MADN system are shown in Fig.~\ref{fig:mle_MADN_withE} together with the indicated confidence intervals obtained from Monte-Carlo sampling. Black dots at the $x$-axes are the data from the explicit calculations for the six \ahfx values discussed before. Considering first the three distributions with single property sampling in panels (a)-(c), we find the confidence intervals for $\log_{10}(\tau)$ of $[-8.0,-7.4]$ for uncertain $\lambda$, $[-8.0,-6.5]$ for uncertain $E$, and $[-7.9,-7.7]$ for uncertain $J$, respectively. Interestingly, all three share very similar lower limits of the confidence interval, while the upper limits and the shape of the distributions differ significantly. While $P_\lambda(\log_{10}(\tau))$ and $P_J(\log_{10}(\tau))$ appear relatively symmetric, indictaing possible a log-normal distribution of $\tau$, $P_E(\log_{10}(\tau))$ is distinctly asymmetric which might point to a normal distribution of $\tau$ for uncertain energies. From the three distributions and the confidence intervals one can see that the effect of the uncertainty in the coupling elements on the time-of-flight is small (factor of 2 on $\tau$). The effect of uncertainty in the reorganization energy is only slightly higher (factor of four). Uncertainty in the site energies, in contrast, has a much bigger impact on the estimates of $\tau$. From the distribution of $P_\mathbf{x}(\log_{10}(\tau))$ in Fig.~\ref{fig:mle_MADN_withE}(d), one can see than when all parameters are sampled, the result is dominated by the effects from the uncertainty in the site energies, with an overall similar shape and a confidence interval of $[-8.3,-6.6]$, corresponding to a range from about \unit[5]{ns} to about \unit[250]{ns} in $\tau$.

>>>>

Uncertainty quantification (UQ) is concept from computational science which allows for an estimation of confidence intervals for a quantity of interest (QoI) and an analysis of its sensitivity in models of a physical system that contain uncertain, maybe empirical, or noisy, parameters~\cite{sarkar_uncertainty_2017,oconnor_quantifying_2024, chernatynskiy_uncertainty_2013, suleimenova_tutorial_2021,coveney_reliability_2021, coveney_when_2021}. Many common UQ studies focus on models with (partial) differential equations, e.g., drift-diffusion equations in which the diffusivity as {\color{red}[Should this be: servers as the]} parameter, and assume a certain distribution for the values of the parameter(s). For the multiscale model of charge transport, it is not straightforward to cast the large variety of available exchange-correlation functionals into the role of a model parameter with some distribution. To keep the problem tractable, we focus here instead specifically on the exchange part in hybrid functionals~\cite{perdew_rationale_1996,marques_densitybased_2011}, in which a DFT model for the exchange is mixed using a weighting factor $\alpha_\text{HFX}$ with a Hartree--Fock type exchange, i.e.,
%
\begin{equation}
  E_\text{x} = \alpha_\text{HFX} E_\text{x}^\text{HF} + (1-\alpha_\text{HFX})E_\text{x}^\text{DFT}.
  \label{equ:hybrid}
\end{equation}
%
\change{This choice is motivated by the observation that the simulated charge-carrier time-of-flight (ToF) for a set of commonly used functionals --- PBE~\cite{PhysRevLett.77.3865}, B3LYP~\cite{10.1063/1.464913}, PBE0~\cite{adamo_toward_1999}, BHLYP~\cite{10.1063/1.464913,lee_development_1988}, M06L~\cite{10.1063/1.2370993}, TPSS~\cite{PhysRevLett.91.146401} --- (see Tab.~\ref{tab:actual_functionals} in the Appendix) is more sensitive to the amount of included Hartree--Fock exchange than to variations in the generalized-gradient approximation (GGA) parametrization. While these functionals vary in more details, the data shows that for functionals with varying \ahfx, the simulated ToF spans two orders of magnitude, whereas among non-hybrid (meta-)GGA functionals the ToF results {\color{red} are} on the same order of magnitude.} More specifically, we take as the basis the PBE0 functional and scrutinize (i) how the predictions of the multiscale model of charge transport are affected by variation of $\alpha_\text{HFX}$ as a proxy for uncertainties in the choice of DFT functionals, (ii) what the level of confidence is in quantitative predictions, and (iii) what are the most sensitive quantities in the model. In this sense, Eq.~\ref{equ:hybrid} is deceptively simple. For each value of $\alpha_\text{HFX}$, the graph $G$ is constructed using the respective value of the reorganization energy, the $N_\text{mol}$ site energies, and the $N_\text{pair}$ coupling elements, and the dimensionality of the problem from the perspective of UQ is $N_\text{UQ}=1+N_\text{mol}+N_\text{pair}$, which can easily be on the order of $10^{4}-10^{5}$. We consider the simulated time-of-flight (ToF) and the associated mobility the QoI in the following (see Fig.~\ref{fig:MSM}(c)) which then are subject to uncertainties in these $N_\text{UQ}$ parameters, {\em stemming} from the variation in $\alpha_\text{HFX}$. As a prototypical system, we will study hole transport in an amorphous morphology of 2-methyl-9, 10-bis(naphthalen-2-yl)anthracene (MADN), a wide-gap semiconductor that is used extensively as an ambipolar host material in organic light-emitting diodes~\cite{ko_accurate_2019, chang_great_2017}.

@@

The four respective distributions of $\tau$ in the MADN system are shown in Fig.~\ref{fig:mle_MADN_withE} together with the indicated confidence intervals obtained from Monte-Carlo sampling. Black dots at the $x$-axes are the data from the explicit calculations for the six \ahfx values discussed before. Considering first the three distributions with single property sampling in panels (a)-(c), we find the confidence intervals for $\log_{10}(\tau)$ of $[-8.0,-7.4]$ for uncertain $\lambda$, $[-8.0,-6.5]$ for uncertain $E$, and $[-7.9,-7.7]$ for uncertain $J$, respectively. Interestingly, all three share very similar lower limits of the confidence interval, while the upper limits and the shape of the distributions differ significantly. While $P_\lambda(\log_{10}(\tau))$ and $P_J(\log_{10}(\tau))$ appear relatively symmetric, indictaing possible a log-normal distribution of $\tau$, $P_E(\log_{10}(\tau))$ is distinctly asymmetric which might point to a normal distribution of $\tau$ for uncertain energies. From the three distributions and the confidence intervals one can see that the effect of the uncertainty in the coupling elements on the time-of-flight is small (factor of 2 on $\tau$). The effect of uncertainty in the reorganization energy is only slightly higher (factor of four). Uncertainty in the site energies, in contrast, has a much bigger impact on the estimates of $\tau$. From the distribution of $P_\mathbf{x}(\log_{10}(\tau))$ in Fig.~\ref{fig:mle_MADN_withE}(d), one can see that when all parameters are sampled, the result is dominated by the effects from the uncertainty in the site energies, with an overall similar shape and a confidence interval of $[-8.3,-6.6]$, corresponding to a range from about \unit[5]{ns} to about \unit[250]{ns} in $\tau$.
